---
title: "L04 Maps"
subtitle: "Data Visualization (STAT 302)" 
author: "Yizhe Zhang"
pagetitle: "L04 Yizhe Zhang"

format:
  html:
    toc: true
    toc-location: left
    embed-resources: true
    link-external-newwindow: true
    code-fold: show
    fig-dpi: 300
    
execute:
  warning: false

from: markdown+emoji
reference-location: margin
citation-location: margin 
---



## Github Repo Link



https://github.com/yzh349/Stat302.git



## Overview

The goal of this lab is to explore various ways of building maps with `ggplot2`.

Challenges are not mandatory for students to complete. We highly recommend students attempt them though. *We would expect graduate students to attempt the challenges.*

## Datasets

We'll be using the `US_income.rda` dataset which should be placed in the `data/` subdirectory in our **L04** project. You'll also be downloading your own data to build maps.

```{r} 
#| label: load-packages

# Load package(s)
library(tidyverse)
library(sf)
library(tidycensus)
library(mapview)
library(tigris)
library(geodata)
library(raster)
library(janitor)
library(maps)
library(viridis)
library(statebins)
#resolve conflicted functions
conflicted::conflict_prefer("select","dplyr")
conflicted::conflict_prefer("filter","dplyr")

# Load dataset(s)
load(file = "US_income.rda")
```

## Exercise 1

### Plot 1

Make a county map of a US state using `geom_polygon()`. Maybe use your home state or a favorite state. Please do NOT use the state in the ggplot2 book example. 

Optional: Consider adding major cities (or your home town).

*Hints:*

-   See section [6.1](https://ggplot2-book.org/maps.html#polygonmaps) in our book.
-   Void theme

::: {.callout-tip icon="false"}
## Solution

```{r}
ny_counties <- map_data("county" , "new york") %>%
  select(lon = long, lat, group, id = subregion)
  ggplot(ny_counties, aes(lon, lat, group = group)) +
  geom_polygon(fill = NA, color = "grey50") +
  coord_quickmap() +
  ggtitle("NewYork") +
  theme_void()
```

### Plot 2

Now use `geom_sf()` instead. You'll need to download data for this. You can use either the `tigris` ([github page](https://github.com/walkerke/tigris)) or `geodata` packages. Either `tigris`' `counties()` with `cb = TRUE` or `geodata`'s `gadm()` could be useful.

::: {.callout-tip icon="false"}
## Solution

```{r}
usa_dat <- gadm(country = "USA" , level = 2, path = tempdir()) %>%
  st_as_sf()%>%
  clean_names()
ny_data <- usa_dat %>%
  filter(name_1 == "New York")
ggplot(ny_data) +
  geom_sf(fill = NA, color = "grey50") +
  theme_void() + 
  ggtitle("New York")

```

## Exercise 2

Using the `US_income` dataset, recreate the following graphics as precisely as possible.

```{r}
#| label: ex-2-data-wrangle


# Setting income levels
US_income <- mutate(
  US_income,
  income_bins = cut(
    ifelse(is.na(median_income), 25000, median_income),
    breaks = c(0, 40000, 50000, 60000, 70000, 80000),
    labels = c("< $40k", "$40k to $50k", 
               "$50k to $60k", "$60k to $70k", "> $70k"),
    right = FALSE
  )
)
```

### Plot 1

*Hints:*

-   `geom_sf()` --- boundary color is `"grey80"` and size is 0.2
-   `viridis` package (`discrete = TRUE` in `scale_*` function)
-   Void theme

::: {.callout-tip icon="false"}
## Solution

```{r}
ggplot(US_income, aes(fill = income_bins, geometry = geometry)) + geom_sf(color = "grey80" , size = 0.2) + 
  theme_void() +
  scale_fill_viridis(discrete = TRUE) +
  labs(fill = "median\nincome")

```
### Plot 2

*Hints:*

-   `statebins::geom_statebins()`
-   `viridis` package (`discrete = TRUE` in `scale_*` function)
-   Statebins theme

::: {.callout-tip icon="false"}
```{r}
ggplot(US_income, aes(state = name, fill = income_bins)) +
  geom_statebins() +
  scale_fill_viridis(discrete = TRUE) +
  theme_statebins() +
  labs(fill = "median\nincome")

```


## Exercise 3

Pick any city or foreign country to build a map for. You can dress it up or make it as basic as you want. Also welcome to try building a graphic like that depicted at the end of section [6.5](https://ggplot2-book.org/maps.html#rastermaps) --- use a different region though.

::: {.callout-tip icon="false"}
## Solution
```{r}
chn <- gadm(country = "CHN" , level = 1, path = tempdir()) %>%
  st_as_sf()
ggplot(chn) +
  geom_sf(fill = NA, color = "grey50") +
  theme_void() +
  ggtitle("Map of China")
    
```

## Challenge(s)

::: callout-important

<!-- Graduate students are required to complete this challenge. **It is optional for undergraduate students.** -->

It is optional for everyone, but we highly encourage students to give it a try. Several students in past versions of this course used mapping data like this for their final project. Usually using the `leaflet` or `mapview` packages. This one is done with `mapview`, but as an additional challenge students could try doing this with `leaflet` too.

:::

Using the `tidycensus` package and few others, try to create a map like below using [these directions](https://walkerke.github.io/tidycensus/articles/spatial-data.html). Try using a different geographical area and a different variable from the ACS.

::: {.callout-tip icon="false"}
## Solution

YOUR SOLUTION HERE

:::

